Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-05-14 manuscript.tex
cover letter and citations

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color}
\usepackage{CJK}
\usepackage{framed}

@@

Over the last couple of decades, DNA has proven to be much more than a vessel for genetic information. From sensing to computing to directed self-assembly, the programmable and predictable nature of DNA has unlocked numerous unforeseen nanotechnology applications.\citep{Seeman2017DNANanotechnology, Adleman1994MolecularProblems, Rothemund2006FoldingPatterns, Gu2010ALine} Recently, single molecule localization techniques have exploited the rapid and transient binding of short DNA oligomers in order to achieve super-resolution microscopy and optical multiplexing. \cite{Schnitzbauer2017Super-resolutionDNA-PAINT, Shah2019, Strauss2020UpDNA-PAINT} Predictive understanding of the sequence-dependent thermodynamics of DNA hybridization/dehybridization -- the assembly/disassembly of a DNA duplex from two single strands -- has underpinned the rational design of DNA oligomer sequences for nanotechnological applications, where sequence-dependent nearest-neighbor models can accurately account for mismatched pairs, dangling ends, and other non-native bonding effects.\citep{SantaLucia1998AThermodynamics, Santalucia2004TM} Secondary DNA structures such as hairpins and G-quadruplexes have also been studied in depth and leveraged for nanotechnology applications.\citep{Tsukanov2013DetailedOrigami, Mosayebi2014TheFormation, Mergny2019DNANanotechnology} Predictive models of the \textit{dynamical}, as opposed to purely thermodynamical, behaviors of DNA have become increasingly important in developing technologies such as DNA-PAINT (DNA Points Accumulation For Imaging In Nanoscale Topography), but developments in dynamic DNA nanotechnologies have outpaced our understanding of the dynamics themselves. \citep{Deluca2020DynamicDevices, Cordes2010SensingTransfer, Naimark2020DNADevelopment} Many experimental and computational studies have investigated DNA dynamical phenomena from the picosecond to millisecond range.\citep{Yin2011KineticsHybridization, Xiao2019, Hinckley2014Coarse-grainedEffects, Sanstead2016, Porschke1971CooperativeTransition} Kinetic models have been developed for particular DNA processes such as toehold exchanges and optical barcoding \citep{Zhang2009ControlExchange, Shah2019ProgrammingFingerprinting} and supervised machine learning techniques have been combined with experimental measurements to predict the on/off rates as a function of sequence. \citep{Schickinger2018TetheredHelices, Zhang2018PredictingSequence, Shah2019}. A comprehensive understanding of the full dynamical landscape of hybridization/dehybridization accounting for the sequence-dependent metastable states and association/dissociation pathways remains lacking and fundamental questions remain unresolved. For example, it remains unclear the extent to which hybridization of short DNA oligomers largely proceeds in a conventionally assumed ``all-or-nothing'' fashion or if long-lived metastable states facilitate the transition.\citep{Araque2016LatticeCooperativity, Sikora2013ModelingIntermediates, Wyer2014KineticsAT-tracts, Sanstead2016, Sanstead2018DirectDehybridization, Xiao2019} Out-of-register ``shifted'' base paired structures \citep{Phys2014, Maciejczyk2014DNAModel, Araque2016LatticeCooperativity, Xiao2019} and frayed structures \citep{Zgarbova2014BaseRNA, Nonin1995TerminalFraying, Nikolova2012ProbingSimulations, Andreatta2006UltrafastHelix} stand as candidates for metastable states with the potential to mediate substantial deviations from all-or-nothing behavior, but the extent to which these states are kinetically relevant is difficult to determine experimentally and is likely to be highly sequence-dependent. The development of predictive models and design rules with which to engineer DNA strands with tailored hybridization/dehybridization kinetics and pathways is vital to advancing rational design of DNA strands for nanotechnology applications and is also of importance in understanding fundamental biological processes such as transcription and gene regulation.

Our understanding of hybridization dynamics has been built from decades of experiments -- such as temperature-jump, salt-jump, pH-jump, and other perturbative methods -- that rapidly stimulate DNA and monitor relaxation to a new equilibrium.\citep{Morrison1993SensitiveSolution, Wetmur1968KineticsDNA, Craig1971RelaxationOligonucleotides, Porschke1973ThermodynamicsPairs, Williams1989LaserDGCATGC, Narayanan2012ExploringMixing, Chen2007InfluenceHybridization, Sanstead2018DirectDehybridization} More recently, single molecule diffusion and tethered multifluorophore assays have facilitated equilibrium analysis, but these results can be hampered by slow data collection rates and fluorescent tags effects on strand dynamics, particularly for shorter oligomers.\citep{Liu20173DSolution,  Schickinger2018TetheredHelices, Chen2008Base-by-baseSpectroscopy, Dupuis2013Single-moleculeHelices} A number of computational modeling approaches have also been employed to provide molecular-level resolution of hybridization. Simplified lattice models can recapitulate the essential aspects of the hybridization pathways but lack the realism of continuous space representations \citep{Araque2016LatticeCooperativity, Phys2019}. The long time scales associated with hybridization/dehybridization events place them outside the reach of unbiased all-atom molecular dynamics simulations,\citep{Phys2014} but they can be observed by employing enhanced sampling techniques such as metadynamics, umbrella sampling,  transition path sampling, and forward flux sampling,\citep{Piana2007AtomisticTransition, Zerze2021ThermodynamicsSimulations, Hinckley2013AnHybridization, Schmitt2013ExploringSurface, Sambriski2009, Sambriski2009SequencePathways, Hoefert2011MolecularOligonucleotides,Romano2013DNADependence, Hinckley2014Coarse-grainedEffects} or by using elevated temperature or denaturing solvent concentrations to induce one-way dissociation events.\citep{Wong2008TheSimulations, Perez2010Real-timeUnfolding} The effect of the applied bias upon the thermodynamics can be rigorously corrected for using standard reweighting techniques, but approaches to rigorously correct the kinetics, particularly under the conditions of high bias necessary for good sampling, are in their infancy.~\citep{Prinz2011OptimalDynamics, Chodera2011DynamicalTemperatures, Stelzl2017DynamicSimulations, Donati2017GirsanovModels, Donati2018GirsanovSimulations, Quer2018ANCOORDINATES} A number of coarse-grained DNA force fields have been developed that enable direct observation of these events over microsecond time scales via unbiased coarse-grained molecular dynamics (CGMD) simulations, \citep{Romano2013DNADependence, Hinckley2013AnHybridization, Maciejczyk2014DNAModel, Markegard2015, Dans2016MultiscaleDNA} which, up to a speedup factor associated with the smoothing of the underlying free energy landscape inherent to the coarse-graining procedure, can preserve a faithful model of the unbiased dynamics and associated pathways. These models have previously been used to study biological phenomena such as nucleosome dynamics\citep{Lequieu2016Tension-dependentUnwrapping, Lequieu2017InSliding} and transcription factor binding \cite{Terakawa2015P53Simulations, Tan2018DynamicData} as well as nanotechnology applications such as strand displacement \citep{Srinivas2013OnDisplacement, Haley2020DesignDisplacement} and DNA origami \citep{Snodin2019Coarse-grainedOrigami, Doye2020TheOrigami}.

In this work, we study a family of self-complementary pairs of 10-mer DNA oligomers using coarse-grained molecular simulation, machine learning, and data-driven inference of long-time kinetic models to establish new understanding of the influence of sequence upon hybridization/dehybridization kinetics and mechanisms. This family of oligomers -- 5$^\prime$-ATATATATAT-3$^\prime$ (AT-all), 5$^\prime$-GATATATATC-3$^\prime$ (GC-end), 5$^\prime$-ATATGCATAT-3$^\prime$ (GC-core), and 5$^\prime$-ATGATATCAT-3$^\prime$ (GC-mix) -- was designed to probe the influence of the placement of a G:C pair within an otherwise repetitive A:T sequence and has been the subject of our prior experimental investigations using temperature-jump infrared spectroscopy and simple lattice models. \citep{Sanstead2016} We validate the new computational models of hybridization/dehybridization dynamics developed in this work against new and existing experimental data and reinterpret our prior experimental observations in light of the new computational understanding. Consistent with previous studies, \citep{Phys2014,Romano2013DNADependence,Araque2016LatticeCooperativity} we find the degree of repetitiveness in the sequence -- and therefore the kinetic accessibility and thermodynamic stability of non-native states -- leads to richer dynamics populated by a diversity of long-lived metastable states. Our data-driven modeling and analysis rigorously quantifies these behaviors and furnishes accurate predictive models of the hybridization/dehybridization rates, dynamical pathways, and metastable states. Specifically, we demonstrate that disrupting repetitive stretches of A:T bases by placement of interrupting G:C base pairs enables us to tune the landscape from rich six-state to simple two-state ``all-or-nothing'' behavior, and the specific location of the interrupting pair can be used to modulate the stability of long-lived frayed states that facilitate the hybridization/dehybridization process. Analysis of the configurational ensemble within each metastable state reveals the conformational heterogeneity and dynamical interconversions proceeding on time scales that lie below those resolvable within the long-time kinetic model. Taken together, our analyses establish new molecular-level understanding of the sequence-dependent kinetics and pathways through quantitative predictive models for the long-time system dynamics, resolution of the dynamical folding pathways and metastable states, and elementary design rules with which to sculpt and qualitatively alter the dynamical behaviors of the system. We anticipate that this new foundational understanding, and the extension of our approach to more extensive families of DNA sequences, can guide the rational design of DNA oligomers with tailored kinetic properties engineered for DNA nanotechnology applications such as DNA-PAINT. \citep{Shah2019, Strauss2020UpDNA-PAINT}

@@

We performed molecular dynamics simulations of four 10-base self-complementary double-stranded DNA sequences that we have previously studied by ultra-fast temperature-jump infrared spectroscopy \citep{Sanstead2016}: 5$^\prime$-ATATATATAT-3$^\prime$ (AT-all), 5$^\prime$-GATATATATC-3$^\prime$ (GC-end), 5$^\prime$-ATATGCATAT-3$^\prime$ (GC-core), and  5$^\prime$-ATGATATCAT-3$^\prime$ (GC-mix). We modeled the DNA sequences using the coarse-grained 3-Site-Per-Nucleotide v2 (3SPN.2) model that uses three spherical beads to represent the phosphate, deoxyribose sugar, and nitrogenous base of each nucleotide and employs anisotropic interaction potentials to accurately treat intra-strand base-stacking, inter-strand cross-stacking, and base pairing. \citep{Hinckley2013AnHybridization} The model was parameterized against experimental data on bond lengths, bend angles, torsional angles, base step energies, and base stacking free energies, and reliably reproduces the structure, melting temperatures, persistence lengths, and sequence, salt, concentration, and temperature effects on duplex formation. \citep{Hinckley2013AnHybridization} The model enables access to millisecond time scales and has been has widely-adopted to study numerous phenomenon including DNA packing in viral capsids, protein-DNA binding, and nucleosome unwrapping.\citep{Cordoba2017AIons, Lu2020OpenAWSEMSummary, Lequieu2016Tension-dependentUnwrapping} Although the 3SPN.2 model was not directly parameterized against dynamical experimental data, we will show below that the predicted sequence-dependent kinetics and relaxation are, within a corrective scaling factor, in very good agreement with observed experimental trends.

All calculations were performed using the LAMMPS simulation package (\url{http://lammps.sandia.gov}) in accordance with best practices for the 3SPN.2 model. \citep{Plimpton1997Short-RangeDynamics} A single pair of self-complementary sequences were placed in a cubic periodic box with side length 7.8 nm corresponding to a single-strand concentration of 7 mol/L. Solvent effects were modeled implicitly by employing Langevin dynamics \citep{DunwegB.Paul1991BrownianNumbers., Bussi2007AccurateDynamics} with an experimentally motivated per-site friction coefficient of $\xi$ = 9.94 $\times$ 10$^{-11}$ m$^2$/s. \citep{Nkodo2001DiffusionElectrophoresis, Hinckley2013AnHybridization} We specified a 240 mM implicit salt concentration and treated electrostatic interactions using the Debye-H{\"u}ckel with a 5 nm cutoff radius. \citep{Debye1923ZurElektrolyte} Simulations were performed in the NVT ensemble employing a Langevin thermostat. \citep{Schneider1978Molecular-dynamicsTransitions}. Each sequence was simulated at its melting temperature -- AT-all: 309 K, GC-end: 317 K, GC-core: 324 K, GC-mix: 324 K -- in order to maximize the number of spontaneous transitions between dissociated and hybridized states. Melting temperatures for each sequence were determined empirically by comparing the ratio of hybridized to dissociated populations over a 10 K temperature ramp centered on the Nearest Neighbors model predicted melting temperature. The temperature at which the hybridized and dissociated populations were closest was selected. The Langevin equations of motion were integrated using the scheme of Bussi and Parrinello \citep{Bussi2007AccurateDynamics} with a 20 fs integration time step. We performed 40 independent simulations for each of the four sequences with half of the runs initialized from the hybridized state and half from the dissociated state. The initial hybridized state was defined following 3SPN.2 best practices using the crystal structure coordinates of Arnott et al.~\citep{Arnott1976CRCBiology} The dissociated state was generated from the hybridized state by displacing one strand away from the other by 1 nm in each of the $x$, $y$, and $z$ directions. The magnitude of this displacement did not affect the results so long as all native WC bonds were completely broken. Initial bead velocities were assigned from a Maxwell-Boltzmann distribution at the temperature of interest. Each simulation was conducted for 26 $\mu$s and frame saved to disc every 100 ps, requiring $\sim$24 CPU-hours on 28$\times$Intel E5-2680v4 CPU cores. The first 1 $\mu$s of each run was discarded for equilibration providing us with 40$\times$25 $\mu$s = 1 ms of simulation data for each sequence, during which time we observed 55-100 hybridization/dehybridization events.

@@

Markov state models (MSMs) are a powerful approach to infer long-time kinetic models from short molecular simulation trajectories \citep{Sengupta2019AutomatedSelf-assembly} \citep{ Pande2010EverythingAsk, Chodera2014MarkovDynamics, Husic2018MarkovScience, Wehmeyer2019IntroductionV1.0} that we employ to construct high-resolution sequence-dependent kinetic models of DNA hybridization and dissociation. MSMs have recently been implemented to study the hybridization mechanism of one particular 14-mer DNA oligomer, but determining the sequence-dependent kinetics and mechanisms was not the focus of this study.\citep{Jin2019} An energy disconnectivity graph-based approach was used to interrogate the differences in hybridization rates and mechanisms between GGGGGG and GCGCGC hexamers to reveal strong deviations from ``all-or-nothing'' behaviors and the importance of zippering and slithering mechanisms.~\citep{Xiao2019} Kinetic models were constructed not from the dynamical trajectories of the molecular model, but by estimating rate constants between local minima were estimated using a transition state theory approximation. Finally, a recent application of MSMs to the long-time dynamics of short RNA oligonucleotides revealed stacking time scales to be highly sequence dependent.~\citep{Pinamonti2017} In this work, MSMs were constructed for each of the four DNA sequences from the 40$\times$25 $\mu$s simulation trajectories following the six-step protocol detailed in Ref.~\citep{Sidky}: (i) trajectory featurization, (ii) dimensionality reduction, (iii) microstate clustering and microstate transition matrix inference, (iv) macrostate clustering and macrostate transition matrix inference. Calculations were performed using the PyEMMA software package. \citep{Scherer2015PyEMMAModels}

@@

\textbf{Macrostate clustering.} The 200 microstates comprising each system were finally coarsened into our terminal macrostate MSM. The microstate transition matrix for each system was computed at a range of lag times $\tau$ and then diagonalized to recover the corresponding eigenvalues $\lambda_i$ and associated implied time scales $t_i = -\tau / \ln \left| \lambda_i \right|$.~\citep{Wehmeyer2019IntroductionV1.0} The implied time scale plots for the four DNA sequences are presented in \blauw{Fig.~S2}. We observe rapid convergence of the implied time scales $t_i$ with lag time $\tau$ for all systems, motivating the construction of high resolution macrostate MSMs at a lag time $\tau$ = 1.2 ns. For this choice of lag time, we recover 5, 4, 2, and 1 implied time scales for the AT-all, GC-end, GC-core, and GC-mix systems, respectively. The identification of $(i-1)$ implied time scales implies the presence of $(i-1)$ slow modes and motivates the coarsening of the system into $i$ macrostates. We estimate these $i$ macrostates by applying PCCA+ spectral clustering to the leading $(i-1)$ eigenvectors of the microstate transition matrix.~\citep{Roblitz2013FuzzyClassification, Weber2018ImplicationsSimulation, Kube2007AConformations} We then estimate the corresponding 6, 5, 3, and 2 macrostate transition matrices $\mathbf{P}$ for the AT-all, GC-end, GC-core, and GC-mix systems, respectively, by projecting the MD simulation trajectories into these discrete macrostates. These macrostate MSMs constitute our terminal kinetic models. We validate the Markovian nature of the four MSMs by subjecting them to the Chapman-Kolmogorov (CK) test \citep{Noe2009ConstructingSimulations,Phys2011MarkovValidation,  Wehmeyer2019IntroductionV1.0}. This test asserts that the transition matrix for a Markovian (i.e., memoryless) MSM constructed at a lag time $\tau$ should satisfy the condition $\mathbf{P}(k \tau) = \mathbf{P}^k(\tau)$, which states that $k$ successive applications of the transition matrix constructed at a lag time $\tau$ should be equivalent to a single application of the transition matrix constructed at a lag time $k \tau$. We present in \blauw{Fig.~S3} the CK tests for each DNA sequence to demonstrate that the $\tau$ = 1.2 ns models perform very well in predicting transition probabilities out to $k \tau$ = 7.2 ns, validating the Markovian nature and kinetic validity of the four models.

@@

%\bibliographystyle{unsrt}
\bibliography{references_210505}

%\bibliography{refs_mendeley2}

>>>>

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color}
\usepackage{CJK}
\usepackage{framed}
\usepackage{achemso}

@@

Over the last couple of decades, DNA has proven to be much more than a vessel for genetic information. From sensing to computing to directed self-assembly, the programmable and predictable nature of DNA has unlocked numerous unforeseen nanotechnology applications.\citep{Seeman2017DNANanotechnology, Adleman1994MolecularProblems, Rothemund2006FoldingPatterns, Gu2010ALine} Recently, single molecule localization techniques have exploited the rapid and transient binding of short DNA oligomers in order to achieve super-resolution microscopy and optical multiplexing. \cite{Schnitzbauer2017Super-resolutionDNA-PAINT, Shah2019, Strauss2020UpDNA-PAINT} Predictive understanding of the sequence-dependent thermodynamics of DNA hybridization/dehybridization -- the assembly/disassembly of a DNA duplex from two single strands -- has underpinned the rational design of DNA oligomer sequences for nanotechnological applications, where sequence-dependent nearest-neighbor models can accurately account for mismatched pairs, dangling ends, and other non-native bonding effects.\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} Secondary DNA structures such as hairpins and G-quadruplexes have also been studied in depth and leveraged for nanotechnology applications.\citep{Tsukanov2013DetailedOrigami, Mosayebi2014TheFormation, Mergny2019DNANanotechnology} Predictive models of the \textit{dynamical}, as opposed to purely thermodynamical, behaviors of DNA have become increasingly important in developing technologies such as DNA-PAINT (DNA Points Accumulation For Imaging In Nanoscale Topography), but developments in dynamic DNA nanotechnologies have outpaced our understanding of the dynamics themselves. \citep{Deluca2020DynamicDevices, Cordes2010SensingTransfer, Naimark2020DNADevelopment} Many experimental and computational studies have investigated DNA dynamical phenomena from the picosecond to millisecond range.\citep{Yin2011KineticsHybridization, Xiao2019, Hinckley2014Coarse-grainedEffects, Sanstead2016, Porschke1971CooperativeTransition} Kinetic models have been developed for particular DNA processes such as toehold exchanges and optical barcoding \citep{Zhang2009ControlExchange, Shah2019ProgrammingFingerprinting} and supervised machine learning techniques have been combined with experimental measurements to predict the on/off rates as a function of sequence. \citep{Schickinger2018TetheredHelices, Zhang2018PredictingSequence, Shah2019}. A comprehensive understanding of the full dynamical landscape of hybridization/dehybridization accounting for the sequence-dependent metastable states and association/dissociation pathways remains lacking and fundamental questions remain unresolved. For example, it remains unclear the extent to which hybridization of short DNA oligomers largely proceeds in a conventionally assumed ``all-or-nothing'' fashion or if long-lived metastable states facilitate the transition.\citep{Araque2016LatticeCooperativity, Sikora2013ModelingIntermediates, Wyer2014KineticsAT-tracts, Sanstead2016, Sanstead2018DirectDehybridization, Xiao2019} Out-of-register ``shifted'' base paired structures \citep{Hinckley2014Coarse-grainedEffects, Maciejczyk2014DNAModel, Araque2016LatticeCooperativity, Xiao2019} and frayed structures \citep{Zgarbova2014BaseRNA, Nonin1995TerminalFraying, Nikolova2012ProbingSimulations, Andreatta2006UltrafastHelix} stand as candidates for metastable states with the potential to mediate substantial deviations from all-or-nothing behavior, but the extent to which these states are kinetically relevant is difficult to determine experimentally and is likely to be highly sequence-dependent. The development of predictive models and design rules with which to engineer DNA strands with tailored hybridization/dehybridization kinetics and pathways is vital to advancing rational design of DNA strands for nanotechnology applications and is also of importance in understanding fundamental biological processes such as transcription and gene regulation.

Our understanding of hybridization dynamics has been built from decades of experiments -- such as temperature-jump, salt-jump, pH-jump, and other perturbative methods -- that rapidly stimulate DNA and monitor relaxation to a new equilibrium.\citep{Morrison1993SensitiveSolution, Wetmur1968KineticsDNA, Craig1971RelaxationOligonucleotides, Porschke1973ThermodynamicsPairs, Williams1989LaserDGCATGC, Narayanan2012ExploringMixing, Chen2007InfluenceHybridization, Sanstead2018DirectDehybridization} More recently, single molecule diffusion and tethered multifluorophore assays have facilitated equilibrium analysis, but these results can be hampered by slow data collection rates and fluorescent tags effects on strand dynamics, particularly for shorter oligomers.\citep{Liu20173DSolution,  Schickinger2018TetheredHelices, Chen2008Base-by-baseSpectroscopy, Dupuis2013Single-moleculeHelices} A number of computational modeling approaches have also been employed to provide molecular-level resolution of hybridization. Simplified lattice models can recapitulate the essential aspects of the hybridization pathways but lack the realism of continuous space representations \citep{Araque2016LatticeCooperativity, Phys2019}. The long time scales associated with hybridization/dehybridization events place them outside the reach of unbiased all-atom molecular dynamics simulations,\citep{Hinckley2014Coarse-grainedEffects} but they can be observed by employing enhanced sampling techniques such as metadynamics, umbrella sampling,  transition path sampling, and forward flux sampling,\citep{Piana2007AtomisticTransition, Zerze2021ThermodynamicsSimulations, Hinckley2013AnHybridization, Schmitt2013ExploringSurface, Sambriski2009, Sambriski2009SequencePathways, Hoefert2011MolecularOligonucleotides,Romano2013DNADependence, Hinckley2014Coarse-grainedEffects} or by using elevated temperature or denaturing solvent concentrations to induce one-way dissociation events.\citep{Wong2008TheSimulations, Perez2010Real-timeUnfolding} The effect of the applied bias upon the thermodynamics can be rigorously corrected for using standard reweighting techniques, but approaches to rigorously correct the kinetics, particularly under the conditions of high bias necessary for good sampling, are in their infancy.~\citep{Prinz2011OptimalDynamics, Chodera2011DynamicalTemperatures, Stelzl2017DynamicSimulations, Donati2017GirsanovModels, Donati2018GirsanovSimulations, Quer2018ANCOORDINATES} A number of coarse-grained DNA force fields have been developed that enable direct observation of these events over microsecond time scales via unbiased coarse-grained molecular dynamics (CGMD) simulations, \citep{Romano2013DNADependence, Hinckley2013AnHybridization, Maciejczyk2014DNAModel, Markegard2015, Dans2016MultiscaleDNA} which, up to a speedup factor associated with the smoothing of the underlying free energy landscape inherent to the coarse-graining procedure, can preserve a faithful model of the unbiased dynamics and associated pathways. These models have previously been used to study biological phenomena such as nucleosome dynamics\citep{Lequieu2016Tension-dependentUnwrapping, Lequieu2017InSliding} and transcription factor binding \cite{Terakawa2015P53Simulations, Tan2018DynamicData} as well as nanotechnology applications such as strand displacement \citep{Srinivas2013OnDisplacement, Haley2020DesignDisplacement} and DNA origami \citep{Snodin2019Coarse-grainedOrigami, Doye2020TheOrigami}.

In this work, we study a family of self-complementary pairs of 10-mer DNA oligomers using coarse-grained molecular simulation, machine learning, and data-driven inference of long-time kinetic models to establish new understanding of the influence of sequence upon hybridization/dehybridization kinetics and mechanisms. This family of oligomers -- 5$^\prime$-ATATATATAT-3$^\prime$ (AT-all), 5$^\prime$-GATATATATC-3$^\prime$ (GC-end), 5$^\prime$-ATATGCATAT-3$^\prime$ (GC-core), and 5$^\prime$-ATGATATCAT-3$^\prime$ (GC-mix) -- was designed to probe the influence of the placement of a G:C pair within an otherwise repetitive A:T sequence and has been the subject of our prior experimental investigations using temperature-jump infrared spectroscopy and simple lattice models. \citep{Sanstead2016} We validate the new computational models of hybridization/dehybridization dynamics developed in this work against new and existing experimental data and reinterpret our prior experimental observations in light of the new computational understanding. Consistent with previous studies, \citep{Hinckley2014Coarse-grainedEffects,Romano2013DNADependence,Araque2016LatticeCooperativity} we find the degree of repetitiveness in the sequence -- and therefore the kinetic accessibility and thermodynamic stability of non-native states -- leads to richer dynamics populated by a diversity of long-lived metastable states. Our data-driven modeling and analysis rigorously quantifies these behaviors and furnishes accurate predictive models of the hybridization/dehybridization rates, dynamical pathways, and metastable states. Specifically, we demonstrate that disrupting repetitive stretches of A:T bases by placement of interrupting G:C base pairs enables us to tune the landscape from rich six-state to simple two-state ``all-or-nothing'' behavior, and the specific location of the interrupting pair can be used to modulate the stability of long-lived frayed states that facilitate the hybridization/dehybridization process. Analysis of the configurational ensemble within each metastable state reveals the conformational heterogeneity and dynamical interconversions proceeding on time scales that lie below those resolvable within the long-time kinetic model. Taken together, our analyses establish new molecular-level understanding of the sequence-dependent kinetics and pathways through quantitative predictive models for the long-time system dynamics, resolution of the dynamical folding pathways and metastable states, and elementary design rules with which to sculpt and qualitatively alter the dynamical behaviors of the system. We anticipate that this new foundational understanding, and the extension of our approach to more extensive families of DNA sequences, can guide the rational design of DNA oligomers with tailored kinetic properties engineered for DNA nanotechnology applications such as DNA-PAINT. \citep{Shah2019, Strauss2020UpDNA-PAINT}

@@

We performed molecular dynamics simulations of four 10-base self-complementary double-stranded DNA sequences that we have previously studied by ultra-fast temperature-jump infrared spectroscopy \citep{Sanstead2016}: 5$^\prime$-ATATATATAT-3$^\prime$ (AT-all), 5$^\prime$-GATATATATC-3$^\prime$ (GC-end), 5$^\prime$-ATATGCATAT-3$^\prime$ (GC-core), and  5$^\prime$-ATGATATCAT-3$^\prime$ (GC-mix). We modeled the DNA sequences using the coarse-grained 3-Site-Per-Nucleotide v2 (3SPN.2) model that uses three spherical beads to represent the phosphate, deoxyribose sugar, and nitrogenous base of each nucleotide and employs anisotropic interaction potentials to accurately treat intra-strand base-stacking, inter-strand cross-stacking, and base pairing. \citep{Hinckley2013AnHybridization} The model was parameterized against experimental data on bond lengths, bend angles, torsional angles, base step energies, and base stacking free energies, and reliably reproduces the structure, melting temperatures, persistence lengths, and sequence, salt, concentration, and temperature effects on duplex formation. \citep{Hinckley2013AnHybridization} The model enables access to millisecond time scales and has been has widely-adopted to study numerous phenomenon including DNA packing in viral capsids, protein-DNA binding, and nucleosome unwrapping.\citep{Cordoba2017AIons, Lu2020OpenAWSEMSimulations, Lequieu2016Tension-dependentUnwrapping} Although the 3SPN.2 model was not directly parameterized against dynamical experimental data, we will show below that the predicted sequence-dependent kinetics and relaxation are, within a corrective scaling factor, in very good agreement with observed experimental trends.

All calculations were performed using the LAMMPS simulation package (\url{http://lammps.sandia.gov}) in accordance with best practices for the 3SPN.2 model. \citep{Plimpton1997FastDynamics} A single pair of self-complementary sequences were placed in a cubic periodic box with side length 7.8 nm corresponding to a single-strand concentration of 7 mol/L. Solvent effects were modeled implicitly by employing Langevin dynamics \citep{DunwegB.Paul1991BrownianNumbers., Bussi2007AccurateDynamics} with an experimentally motivated per-site friction coefficient of $\xi$ = 9.94 $\times$ 10$^{-11}$ m$^2$/s. \citep{Nkodo2001DiffusionElectrophoresis, Hinckley2013AnHybridization} We specified a 240 mM implicit salt concentration and treated electrostatic interactions using the Debye-H{\"u}ckel with a 5 nm cutoff radius. \citep{Debye1923ZurElektrolyte} Simulations were performed in the NVT ensemble employing a Langevin thermostat. \citep{Schneider1978Molecular-dynamicsTransitions}. Each sequence was simulated at its melting temperature -- AT-all: 309 K, GC-end: 317 K, GC-core: 324 K, GC-mix: 324 K -- in order to maximize the number of spontaneous transitions between dissociated and hybridized states. Melting temperatures for each sequence were determined empirically by comparing the ratio of hybridized to dissociated populations over a 10 K temperature ramp centered on the Nearest Neighbors model predicted melting temperature. The temperature at which the hybridized and dissociated populations were closest was selected. The Langevin equations of motion were integrated using the scheme of Bussi and Parrinello \citep{Bussi2007AccurateDynamics} with a 20 fs integration time step. We performed 40 independent simulations for each of the four sequences with half of the runs initialized from the hybridized state and half from the dissociated state. The initial hybridized state was defined following 3SPN.2 best practices using the crystal structure coordinates of Arnott et al.~\citep{Arnott1976CRCBiology} The dissociated state was generated from the hybridized state by displacing one strand away from the other by 1 nm in each of the $x$, $y$, and $z$ directions. The magnitude of this displacement did not affect the results so long as all native WC bonds were completely broken. Initial bead velocities were assigned from a Maxwell-Boltzmann distribution at the temperature of interest. Each simulation was conducted for 26 $\mu$s and frame saved to disc every 100 ps, requiring $\sim$24 CPU-hours on 28$\times$Intel E5-2680v4 CPU cores. The first 1 $\mu$s of each run was discarded for equilibration providing us with 40$\times$25 $\mu$s = 1 ms of simulation data for each sequence, during which time we observed 55-100 hybridization/dehybridization events.

@@

Markov state models (MSMs) are a powerful approach to infer long-time kinetic models from short molecular simulation trajectories \citep{Sengupta2019AutomatedSelf-assembly} \citep{ Pande2010EverythingAsk, Chodera2014MarkovDynamics, Husic2018MarkovScience, Wehmeyer2019IntroductionSoftware} that we employ to construct high-resolution sequence-dependent kinetic models of DNA hybridization and dissociation. MSMs have recently been implemented to study the hybridization mechanism of one particular 14-mer DNA oligomer, but determining the sequence-dependent kinetics and mechanisms was not the focus of this study.\citep{Jin2019} An energy disconnectivity graph-based approach was used to interrogate the differences in hybridization rates and mechanisms between GGGGGG and GCGCGC hexamers to reveal strong deviations from ``all-or-nothing'' behaviors and the importance of zippering and slithering mechanisms.~\citep{Xiao2019} Kinetic models were constructed not from the dynamical trajectories of the molecular model, but by estimating rate constants between local minima were estimated using a transition state theory approximation. Finally, a recent application of MSMs to the long-time dynamics of short RNA oligonucleotides revealed stacking time scales to be highly sequence dependent.~\citep{Pinamonti2017} In this work, MSMs were constructed for each of the four DNA sequences from the 40$\times$25 $\mu$s simulation trajectories following the six-step protocol detailed in Ref.~\citep{Sidky}: (i) trajectory featurization, (ii) dimensionality reduction, (iii) microstate clustering and microstate transition matrix inference, (iv) macrostate clustering and macrostate transition matrix inference. Calculations were performed using the PyEMMA software package. \citep{Scherer2015PyEMMAModels}

@@

\textbf{Macrostate clustering.} The 200 microstates comprising each system were finally coarsened into our terminal macrostate MSM. The microstate transition matrix for each system was computed at a range of lag times $\tau$ and then diagonalized to recover the corresponding eigenvalues $\lambda_i$ and associated implied time scales $t_i = -\tau / \ln \left| \lambda_i \right|$.~\citep{Wehmeyer2019IntroductionSoftware} The implied time scale plots for the four DNA sequences are presented in \blauw{Fig.~S2}. We observe rapid convergence of the implied time scales $t_i$ with lag time $\tau$ for all systems, motivating the construction of high resolution macrostate MSMs at a lag time $\tau$ = 1.2 ns. For this choice of lag time, we recover 5, 4, 2, and 1 implied time scales for the AT-all, GC-end, GC-core, and GC-mix systems, respectively. The identification of $(i-1)$ implied time scales implies the presence of $(i-1)$ slow modes and motivates the coarsening of the system into $i$ macrostates. We estimate these $i$ macrostates by applying PCCA+ spectral clustering to the leading $(i-1)$ eigenvectors of the microstate transition matrix.~\citep{Roblitz2013FuzzyClassification, Weber2018ImplicationsSimulation, Kube2007AConformations} We then estimate the corresponding 6, 5, 3, and 2 macrostate transition matrices $\mathbf{P}$ for the AT-all, GC-end, GC-core, and GC-mix systems, respectively, by projecting the MD simulation trajectories into these discrete macrostates. These macrostate MSMs constitute our terminal kinetic models. We validate the Markovian nature of the four MSMs by subjecting them to the Chapman-Kolmogorov (CK) test \citep{Noe2009ConstructingSimulations,Prinz2011OptimalDynamics, Wehmeyer2019IntroductionSoftware}. This test asserts that the transition matrix for a Markovian (i.e., memoryless) MSM constructed at a lag time $\tau$ should satisfy the condition $\mathbf{P}(k \tau) = \mathbf{P}^k(\tau)$, which states that $k$ successive applications of the transition matrix constructed at a lag time $\tau$ should be equivalent to a single application of the transition matrix constructed at a lag time $k \tau$. We present in \blauw{Fig.~S3} the CK tests for each DNA sequence to demonstrate that the $\tau$ = 1.2 ns models perform very well in predicting transition probabilities out to $k \tau$ = 7.2 ns, validating the Markovian nature and kinetic validity of the four models.

@@

%\bibliographystyle{unsrt}
%\bibliographystyle{achemso} % should be pre-defined in usepackage
\bibliography{references_210514}
